---
title: "Simulation exercise"
author: "David Marquinez"
output:
  pdf_document: default
  html_document:
    keep_md: yes
---
```{r setup, warning= FALSE , include= FALSE , echo = TRUE}
library(R.utils)
library(ggplot2)
colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color, 
      x)
  } else x
}
```

## Overview

In this analysis we will investigate the exponential distribution and compare it with the Central Limit Theorem. We will set lambda = 0.2 for all of the simulations. We will investigate the distribution of averages of 40 exponentials over a thousand simulations.    

## Simulations  

Setting the variables for lambda, the total number of simulations, the number of exponentials. The seed is setted to 588 in order to make this analysis reproducible.

```{r echo = TRUE}
lambda = 0.2
n = 1000
exp = 40
set.seed(588)
mns = NULL
```  

Simulating the data and calculating the mean of each simulation.

```{r echo = TRUE}
for(i in 1:n) mns = c(mns,mean(rexp(40,lambda)))
```  

## Sample Mean vs Theoretical Mean  

The **sample mean** is  

```{r echo=TRUE}
(sampleMean = mean(mns))
```  

The **theoretical mean** is  

```{r echo=TRUE}
(theoreticalMean=1/lambda)
```  

The **disparity** between the theoretical and the sample mean is  
```{r echo=TRUE}
printf("%.4f %%",100*(1-theoreticalMean/sampleMean))
```

## Sample Variance versus Theoretical Variance  
The **sample variance** is  

```{r echo=TRUE}
(sampleVar = var(mns))
```  

The **theoretical variance** is  

```{r echo=TRUE}
(theoreticalVar=(1/lambda)^2/exp)
``` 
The **disparity between** the theoretical and the sample variance is  
```{r echo=TRUE}
printf("%.4f %%",100*(1-theoreticalVar/sampleVar))
```
## Distribution

```{r echo = TRUE , warning = FALSE}
labelSampleMean <- bquote(Sample~mean==.(sampleMean))
labelTheoreticalMean <- bquote(Theoretical~mean~is~1/lambda==.(theoreticalMean))
#  expression(paste("Theoretical mean=1/",lambda,"=",v.1))
mnsDF <- data.frame(mns)
g <- ggplot(mnsDF,aes(mns))
g + geom_histogram(aes(y=..density..),color="blue",
                   fill="lightblue",bins=30) +
    geom_vline(aes(xintercept=mean(mns)),
               color="blue",size=2,
               linetype="twodash") + 
    annotate(geom = "text", x=mean(mns)*1.35,y=.55,
             label=labelTheoreticalMean, color= "red",
             alpha= 0.7 ) + 
    geom_vline(aes(xintercept=sampleMean),
               color="red",size=2,
               linetype="dotdash") + 
    annotate(geom = "text", x=mean(mns)*1.4,y=.45,
             label=labelSampleMean, color= "blue",
             alpha= 0.7 ) +
    stat_function(fun = dnorm,
            args=list(mean=1/lambda,sd=theoreticalVar^(1/2)),
              color = "red",size=1, alpha=.5) +
    geom_density(color = "blue",size=1,alpha=.5)
```  

As we can see in the plot above, the density of the simulation (`r colorize("blue line","blue")`) is similar to the density of a normal distribution (`r colorize("red line","red")`). Also the means (vertical lines, `r colorize("blue for the sample mean","blue")` and `r colorize("red for the theoretical","red")`) are overlapped because they are pretty similar.