---
title: "Basic Inferential Data Analysis"
author: "David Marquinez"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---
```{r setup, warning= FALSE , include= FALSE , echo = TRUE}
library(R.utils)
library(ggplot2)
colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color, 
      x)
  } else x
}
```

## Overview  

The data we are going to use is a study of the effect of vitamin C on tooth growth in guinea pig.  

## Summary of the data  

We are going to see the structure and a quick view to the data.
```{r echo=TRUE}
data("ToothGrowth")
str(ToothGrowth)
```  
The tooth length 
```{r echo=TRUE}
summary(ToothGrowth$len)
```  
The dose were `r unique(ToothGrowth$dose)` in miligrams/day.  
The supplement type were coded as `r colorize("OJ","orange")`,`r colorize("VC","blue")` and they mean `r colorize("orange juice","orange")` or `r colorize("ascorbic acid","blue")` (a form of vitamin C).  

## Comparation   

```{r echo=TRUE}
levels(ToothGrowth$supp) <- c("Orange Juice", "Ascorbic Acid")
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
g <- ggplot(ToothGrowth, 
            aes(x = dose, y = len, color = supp))
g + facet_grid(. ~ supp) + 
  geom_boxplot( aes(fill = supp),show.legend = F) +
  scale_fill_manual(values = c("orange", "blue")) +
  xlab("Dose (mg/day)") + ylab("Length")
```  

## Hypothesis Tests  

### H0: Both delivery modes of Vitamin C have the same influence on tooth growth. 

```{r echo=TRUE}
len <- ToothGrowth$len
supp <- ToothGrowth$supp
dose <- ToothGrowth$dose
t1 <- t.test(len[supp=="Orange Juice"],
             len[supp=="Ascorbic Acid"],
             paired=F,conf.level = 0.95)
```  

The confidence interval is `r t1$conf.int[1:2]` with a confidence level of `r 100*attr(t1$conf.int,"conf.level")` % and the p-value is `r t1$p.value` greather than 0.05.  
We do not reject the null hypothesis.  

### H1 With a dose of 0.5 mg/day the two delivery methods have the same influence on tooth growth.  
```{r echo=TRUE}
t2 <- t.test(len[supp=="Orange Juice" & dose==0.5] ,
             len[supp=="Ascorbic Acid" & dose==0.5],
             paired=F, conf.level = 0.95)
```  

The confidence interval is `r t2$conf.int[1:2]` with a confidence level of `r 100*attr(t2$conf.int,"conf.level")` % and the p-value is `r t2$p.value` much lower than 0.05.  
We  reject the null hypothesis.  

### H2 With a dose of 1 mg/day the two delivery methods have the same influence on tooth growth.  
```{r echo=TRUE}
t2 <- t.test(len[supp=="Orange Juice" & dose==1] ,
             len[supp=="Ascorbic Acid" & dose==1], 
             paired=F, conf.level = 0.95)
```  

The confidence interval is `r t2$conf.int[1:2]` with a confidence level of `r 100*attr(t2$conf.int,"conf.level")` % and the p-value is `r t2$p.value` much lower than 0.05.  
We  reject the null hypothesis.  

### H3 With a dose of 2 mg/day the two delivery methods have the same influence on tooth growth.  
```{r echo=TRUE}
t2 <- t.test(len[supp=="Orange Juice" & dose==2] ,
             len[supp=="Ascorbic Acid" & dose==2],
             paired=F, conf.level = 0.95)
```  

The confidence interval is `r t2$conf.int[1:2]` with a confidence level of `r 100*attr(t2$conf.int,"conf.level")` % and the p-value is `r t2$p.value` greather than 0.05.  
We  do not reject the null hypothesis.

## Conclusion  
#### **`r colorize("Orange Juice","orange")`** has more influence on tooth growth with a dose < 2 mg/day compared to **`r colorize("Ascorbic Acid","blue")`**.  
#### With a dose of 2 mg/day both delivery methods have the same influence.